---
title: "Hearthstone Arena Statistic Exploration"
output: html_notebook
---

Hearthstone just released some official Arena card data.  I think this is super cool since I don't know they have every fully released this data to the public.  The community has been asking for some official data for some time. 

https://playhearthstone.com/en-gb/blog/21964711/developer-insights-12-0-arena-update-with-kris-zierhut

let's load the data and have a look.

```{r}
library(dplyr)
appearence_data <- read.csv('data/Appearance_Rates.csv')
```

Ok so the first thing we might do is see what the top 10 cards are in each class

```{r}
top_10_in_each_class <- appearence_data %>%
  group_by(Draft.Class) %>%
  top_n(10, Average)
```

So that gives us top 10 cards by average occurances/draft for each class.  I had to look up that function `top_n`.  Let's get a little more infomation on the function.

```{r}
help("top_n")
```

So pretty great function, it also can order by ascending by using a minus `-` flag.

